---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Measuring the Acceptance of Homosexuality Across Countries and Over Time"
thanks: "Corresponding author: [frederick-solt@uiowa.edu](mailto:frederick-solt@uiowa.edu).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Byung-Deuk Woo
  affiliation: Pohang University of Science and Technology
- name: Hyein Ko
  affiliation: University of Iowa
- name: Yuehong Cassandra Tai
  affiliation: Pennsylvania State University
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
anonymous: false
abstract: " [0/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---

\pagenumbering{gobble}

# Authors {.unnumbered}
-   Byung-Deuk Woo, ORCID: <https://orcid.org/0000-0001-6902-7576>, Postdoctoral Researcher, Institute of Social Data Science, Pohang University of Science and Technology, [byungdeukwoo\@gmail.com](mailto:byungdeukwoo@gmail.com){.email}
-   Hyein Ko, ORCID: <https://orcid.org/0000-0002-9497-9656>, Doctoral Candidate, Department of Political Science, University of Iowa, [hyein-ko\@uiowa.edu](mailto:hyein-ko@uiowa.edu){.email}
-   Yuehong Cassandra Tai, ORCID: <https://orcid.org/0000-0001-7303-7443>, Postdoctoral Fellow, Center for Social Data Analytics, Pennsylvania State University, [yhcasstai\@psu.edu](mailto:yhcasstai@psu.edu){.email}
-   Yue Hu, ORCID: <https://orcid.org/0000-0002-2829-3971>, Associate Professor, Department of Political Science, Tsinghua University, [yuehu\@tsinghua.edu.cn](mailto:yuehu@tsinghua.edu.cn){.email}
-   Frederick Solt, ORCID: <https://orcid.org/0000-0002-3154-6132>, Associate Professor, Department of Political Science, University of Iowa, [frederick-solt\@uiowa.edu](mailto:frederick-solt@uiowa.edu){.email}

\pagebreak

```{=tex}
\renewcommand{\baselinestretch}{1}
\selectfont
\maketitle
\renewcommand{\baselinestretch}{1.5}
\selectfont
```

```{=tex}
\begin{abstract}
Lorem ipsum blah blah blah
\end{abstract}
```
\pagebreak

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dpi = 600,
  fig.width = 7,
  fig.height = 4,
  plot = function(x, options)  {
    hook_plot_tex(x, options)
  }
)

# If `DCPOtools` is not yet installed:
# remotes::install_github("fsolt/DCPOtools")

if (!require(pacman)) install.packages("pacman")
library(pacman)
# load all the packages you will use below 
p_load(
  DCPOtools,
  cmdstanr,
  tidyverse,
  here,
  countrycode,
  patchwork,
  ggthemes,
  ggdist,
  imputeTS,
  rsdmx,
  osfr,
  tabulizer,
  brms,
  tidybayes,
  repmis,
  rvest,
  vroom,
  modelsummary,
  kableExtra
) 
```

```{r define_funs}
# define functions
validation_plot <- function(v_data_raw,
                            lab_x = .38, lab_y = 92,
                            theta_summary, theta_results) {
    
    # defaults per https://stackoverflow.com/a/49167744/2620381
    if ("theta_summary" %in% ls(envir = .GlobalEnv) & missing(theta_summary))
        theta_summary <- get("theta_summary", envir = .GlobalEnv)
    if ("theta_results" %in% ls(envir = .GlobalEnv) & missing(theta_results))
        theta_results <- get("theta_results", envir = .GlobalEnv)

    median_val <- Vectorize(function(x) median(1:x),
                            vectorize.args = "x")
    
    v_vars <- v_data_raw %>% 
      select(item0 = item) %>% 
      unique() %>% 
      mutate(v_val = str_extract(item0, "\\d+") %>% 
               as.numeric() %>% 
               median_val(.) %>%
               `+`(if_else(str_detect(item0, "disc3"), .1, .6)) %>% 
               round())
    
    validation_summarized <- v_data_raw %>% 
      DCPOtools::format_dcpo(scale_q = v_vars$item0[[1]], # these arguments are required
                             scale_cp = 1) %>% # but they don't matter
      pluck("data") %>% 
      mutate(item0 = str_remove(item, " \\d or higher"),
             title = factor(title, 
                            levels = v_data_raw %>%
                              pull(title) %>%
                              unique())) %>% 
      right_join(v_vars, by = "item0") %>%
      arrange(title) %>% 
      filter(str_detect(item, paste(v_val, "or higher"))) %>%
      mutate(iso2c = countrycode::countrycode(country,
                                              origin = "country.name",
                                              destination = "iso2c"),
             prop = if_else(neg, 1-y_r/n_r, y_r/n_r),
             se = sqrt((prop*(1-prop))/n),
             prop_90 = prop + qnorm(.9)*se,
             prop_10 = prop - qnorm(.9)*se) %>%
      inner_join(theta_summary %>% select(-kk, -tt), by = c("country", "year"))
    
    validation_cor <- theta_results %>%
      inner_join(validation_summarized %>%
                   select(country, year, title, prop, se),
                 by = c("country", "year")) %>% 
      rowwise() %>% 
      mutate(sim = rnorm(1, mean = prop, sd = se)) %>% 
      ungroup() %>% 
      select(title, theta, sim, draw) %>% 
      nest(data = c(theta, sim)) %>% 
      mutate(r = lapply(data, function(df) cor(df)[2,1]) %>% 
               unlist()) %>%
      select(-data) %>% 
      group_by(title) %>% 
      summarize(r = paste("R =", round(mean(r), 2)))

    if ({validation_summarized %>%
        pull(country) %>%
        unique() %>% 
        length()} > 1) {
      val_plot <- validation_summarized %>%
        ggplot(aes(x = mean,
                   y = prop * 100)) +
        geom_segment(aes(x = q10, xend = q90,
                         y = prop * 100, yend = prop * 100),
                     na.rm = TRUE,
                     alpha = .2) +
        geom_segment(aes(x = mean, xend = mean,
                         y = prop_90 * 100, yend = prop_10 * 100),
                     na.rm = TRUE,
                     alpha = .2) +
        geom_smooth(method = 'lm', formula = 'y ~ x', se = FALSE) +
        facet_wrap(~ title, ncol = 4) +
        geom_label(data = validation_cor, aes(x = lab_x,
                                              y = lab_y,
                                              label = r),
                   size = 2)
    } else {
      val_plot <- validation_summarized %>%
        ggplot(aes(x = year,
                   y = mean)) +
        geom_line() +
        geom_ribbon(aes(ymin = q10,
                        ymax = q90,
                        linetype = NA),
                    alpha = .2) +
        geom_point(aes(y = prop),
                   fill = "black",
                   shape = 21,
                   size = .5,
                   na.rm = TRUE) +
        geom_path(aes(y = prop),
                  linetype = 3,
                  na.rm = TRUE,
                  alpha = .7) +
        geom_segment(aes(x = year, xend = year,
                         y = prop_90, yend = prop_10),
                     na.rm = TRUE,
                     alpha = .2) +
        facet_wrap(~ title, ncol = 4) +
        geom_label(data = validation_cor, aes(x = lab_x,
                                              y = lab_y,
                                              label = r),
                   size = 2)
    }
    
    return(val_plot)
}

covered_share_of_spanned <- function(dcpo_input_raw) {
  n_cy <- dcpo_input_raw %>%
    distinct(country, year) %>% 
    nrow()
  
  spanned_cy <- dcpo_input_raw %>% 
    group_by(country) %>% 
    summarize(years = max(year) - min(year) + 1) %>% 
    summarize(n = sum(years)) %>% 
    pull(n)
  
  {(n_cy/spanned_cy) * 100}
}

get_coef <- function(iv, results_df = coef_data, type = "both", width = .95) {
  result_var <- results_df %>% 
    filter(.width == width) %>% 
    pull(.variable) %>% 
    str_subset(iv)
  
  if (!type=="both") {
    res <- results_df %>% 
      filter(.variable == result_var & .width == width) %>% 
      pull({{type}})
  } else {
    sc <- results_df %>% 
      filter(.variable == result_var & .width == width) %>% 
      pull(std_coef)
    
    ci <- results_df %>% 
      filter(.variable == result_var & .width == width) %>% 
      pull(ci)
    
    res <- paste0(sc, " (95% c.i.: ", ci, ")")
  }
  
  return(res)
}

by2sd <- function(var) {
  dich <- stats::na.omit(unique(var)) %>% 
    sort() %>% identical(c(0, 1))
  if (dich) 
    sd <- 1
  else 
    sd <- 2 * stats::sd(var, na.rm = TRUE)
  
  return(sd)
}

set.seed(324)
```


```{r dcpo_input_raw, include=FALSE, cache = TRUE, cache.extra = tools::md5sum(here::here("data-raw", "surveys_gm.csv"))}
surveys_gm <- read_csv(here::here("data-raw", "surveys_gm.csv"),
                        col_types = "cccc")

dcpo_input_raw <- DCPOtools::dcpo_setup(vars = surveys_gm,
                                        datapath = here::here("..",
                                                              "data", "dcpo_surveys"),
                                        file = here::here("data",
                                                          "dcpo_input_raw.csv"))
```

```{r gm_summary_stats}
dcpo_input_raw <- read_csv(here::here("data", "dcpo_input_raw.csv"),
                                  col_types = "cdcddcd")

process_dcpo_input_raw <- function(dcpo_input_raw_df) {
  dcpo_input_raw_df %>% 
  with_min_yrs(3) %>% 
  with_min_cy(5) %>% 
  group_by(country) %>% 
  mutate(cc_rank = n()) %>% 
  ungroup() %>% 
  arrange(-cc_rank)
} 

dcpo_input_raw1 <- process_dcpo_input_raw(dcpo_input_raw)

n_surveys <- surveys_gm %>%
  distinct(survey) %>% 
  nrow()

n_items <- dcpo_input_raw1 %>%
  distinct(item) %>% 
  nrow()

n_countries <- dcpo_input_raw1 %>%
  distinct(country) %>% 
  nrow()

n_cy <- dcpo_input_raw1 %>%
  distinct(country, year) %>% 
  nrow() %>% 
  scales::comma()

n_years <- as.integer(summary(dcpo_input_raw1$year)[6]-summary(dcpo_input_raw1$year)[1])

spanned_cy <- dcpo_input_raw1 %>% 
  group_by(country) %>% 
  summarize(years = max(year) - min(year) + 1) %>% 
  summarize(n = sum(years)) %>% 
  pull(n) %>% 
  scales::comma()

total_cy <- {n_countries * n_years} %>% 
  scales::comma()

year_range <- paste("from",
                    summary(dcpo_input_raw1$year)[1],
                    "to",
                    summary(dcpo_input_raw1$year)[6])

n_cyi <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>% 
  nrow() %>% 
  scales::comma()

back_to_numeric <- function(string_number) {
  string_number %>% 
    str_replace(",", "") %>% 
    as.numeric()
}

covered_share_of_spanned <- {back_to_numeric(n_cy)/back_to_numeric(spanned_cy) * 100}
```


# Examining the Source Data on Attitudes Toward Homosexuality
Surveys have often included questions about homosexuality over the past half-century, but the resulting data are both sparse and incomparable.
That is, these data are unavailable for many countries and years, and they are generated by many different survey items.
In all, we identified `r n_items` items that were asked in no fewer than five country-years in countries surveyed at least twice; these items were drawn from `r n_surveys` different national and cross-national survey datasets.^[
The complete list of survey items is included in the Appendix.]
Together, these items were asked in `r n_countries` different countries in at least two time points over the `r n_years` years `r year_range`, yielding a total of `r n_cyi` country-year-item observations.
Observations for every year in each country surveyed would total `r total_cy`, and a complete set of country-year-items would include `r {n_countries * n_years * n_items} %>% scales::comma()` observations.
Viewed from this complete-data perspective, the available data can be seen to be very, very sparse.
More optimistically, we do have in the source data `r n_cy` country-years for which we have at least _some_ information about the extent of the acceptance of homosexuality in the population, that is, some `r round(covered_share_of_spanned)`% of the `r spanned_cy` country-years spanned by the data we collected.
Still, the many different survey items employed renders these data incomparable and makes them difficult to use together.


<!-- This should be reworded -->

Observations for every year in each country surveyed would total `r total_cy`, and a complete set of country-year-items would include `r {n_countries * n_years * n_items} %>% scales::comma()` observations.
Viewed from this complete-data perspective, the available data can be seen to be very, very sparse.
More optimistically, we do have in the source data `r n_cy` country-years for which we have at least _some_ information about the extent of the acceptance of homosexuality in the population, that is, some `r round(covered_share_of_spanned)`% of the `r spanned_cy` country-years spanned by the data we collected.
But there can be no denying that the many different survey items employed renders these data incomparable and difficult to use together.

```{r item_and_country_plots, fig.height = 3.5, fig.cap = "Countries and Years with the Most Observations in the Gay Rights Source Data \\label{item_country_plots}"}
items_plot <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(item) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(item, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Country-Years\nObserved") +
  ggtitle("Items")

just10_cy <- dcpo_input_raw1 %>% filter(item == "just10") %>% distinct(country, year) %>% nrow()

just10_surveys <- dcpo_input_raw1 %>% filter(item == "just10") %>% distinct(survey) %>% pull(survey)

countries_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country)) %>% 
  distinct(country, year, item) %>% 
  count(country) %>%
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Year-Items\nObserved") +
  ggtitle("Countries")

cby_plot <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  ggplot(aes(forcats::fct_reorder(country, n, .desc = TRUE), n)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Years\nObserved") +
  ggtitle("Countries")


ybc_plot <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>%
  ggplot(aes(year, nn)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        # axis.text.x  = element_text(angle = 90, vjust = .45, hjust = .95),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(hjust = 0.5, size = 11)) +
  ylab("Countries\nObserved") +
  ggtitle("Years")

us_obs <- dcpo_input_raw1 %>% 
  distinct(country, year, item) %>%
  count(country) %>%
  filter(country == "United States") %>%
  pull(n)

others <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(country) %>%
  arrange(desc(n)) %>%
  slice(2:5) %>%
  pull(country) %>% 
  paste(collapse = ", ") %>% 
  str_replace(", (\\w+)$", ", and \\1")

countries_cp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year, item) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

countries_cbyp <- dcpo_input_raw1 %>%
  mutate(country = if_else(stringr::str_detect(country, "United"),
                           stringr::str_replace(country, "((.).*) ((.).*)", "\\2.\\4."),
                           country),
         country = stringr::str_replace(country, "South", "S.")) %>% 
  distinct(country, year) %>%
  count(country) %>% 
  arrange(desc(n)) %>% 
  head(12) %>% 
  pull(country)

adding <- setdiff(countries_cbyp, countries_cp) %>% 
  knitr::combine_words()

dropping <- setdiff(countries_cp, countries_cbyp) %>% 
  knitr::combine_words()

y_peak_year <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(year)

y_peak_nn <- dcpo_input_raw1 %>%
  distinct(country, year) %>%
  count(year, name = "nn") %>% 
  filter(nn == max(nn)) %>% 
  pull(nn)

data_poorest <- dcpo_input_raw1 %>%
  distinct(country, year, item) %>%
  count(country) %>%
  arrange(n) %>%
  filter(n == 2) %>%
  pull(country) %>% 
  knitr::combine_words()

wordify_numeral <- function(x) setNames(c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten", "eleven", "twelve", "thirteen", "fourteen", "fifteen", "sixteen", " seventeen", "eighteen", "nineteen"), 1:19)[x]

n_data_poorest <- {data_poorest %>%
    str_split(",") %>% 
    first()} %>% 
  length() %>% 
  wordify_numeral()

(countries_plot + cby_plot) / (ybc_plot)
```

Consider the most frequently asked item in the data we collected, which asks respondents whether they think homosexuality "can always be justified, never be justified, or something in between," using a ten-point scale.
Employed by the Asia Barometer, the European Values Survey, the Latinobarómetro, and the World Values Survey, this question was asked in a total of `r just10_cy` different country-years.
That this constitutes only `r {just10_cy*100/(spanned_cy %>% str_replace(",", "") %>% as.numeric())} %>% round()`% of the country-years spanned by our data---and remember, this is the _most common_ survey item---again underscores just how sparse and incomparable the available public opinion data is on this topic.

The upper left panel of Figure&nbsp;\ref{item_country_plots} shows the dozen countries with the highest count of country-year-item observations.
The United States, with `r us_obs` observations, is far and away the best represented country in the source data, followed by `r others`.
At the other end of the spectrum, `r n_data_poorest` countries---`r data_poorest`---have only the minimum two observations required to be included in the source dataset at all.
The upper right panel shows the twelve countries with the most years observed; this group is similar, but with `r adding` joining the list and `r dropping` dropping off.
The bottom panel counts the countries observed in each year and reveals just how few relevant survey items were asked before 1990.
Country coverage reached its peak in `r y_peak_year`, when surveys in `r y_peak_nn` countries included items on homosexuality.
In the next section, we describe how we are able to make use of all of this sparse and incomparable survey data to generate estimates of public opinion that are comparable across countries and years using a latent variable model.

```{r obs_by_cy, fig.height = 9, fig.width = 6.5, fig.cap = "Source Data Observations by Country and Year \\label{obs_by_cy}"}
dcpo_input_raw1 %>% 
  mutate(country = str_replace(country, "’", "'")) %>% 
  distinct(country, year, item, cc_rank) %>% 
  group_by(country, year) %>% 
  summarize(n = n(),
            cc_rank = cc_rank) %>% 
  ungroup() %>% 
  distinct() %>% 
  ggplot(aes(x = year, 
             y = forcats::fct_reorder(country, cc_rank),
             fill = n)) + 
  geom_tile() +
  scale_fill_viridis_b(option = "B",
                       direction = -1,
                       show.limits = TRUE,
                       n.breaks = 7,
                       name = "Observations") +
  labs(x = NULL, y = NULL) +
  scale_x_continuous(breaks=seq(1972, 2020, 4)) +
  scale_y_discrete(position = "right") +
  theme(legend.justification=c(0, 0), 
        legend.position=c(0.01, 0.01),
        axis.text.y  = element_text(size = 7)) 
```


```{r dcpo_input}
dcpo_input <- DCPOtools::format_dcpo(dcpo_input_raw1,
                                     scale_q = "approve4",
                                     scale_cp = 2)
save(dcpo_input, file = here::here("data", "dcpo_input.rda"))
```


```{r dcpo, eval=FALSE, include=FALSE, results=FALSE}
iter <- 1500
 
dcpo <- cmdstan_model("~/Documents/Projects/DCPO/inst/stan/dcpo.stan")

dcpo_output <- dcpo$sample(
   data = dcpo_input[1:13], 
   max_treedepth = 14,
   adapt_delta = 0.99,
   step_size = 0.005,
   seed = 324, 
   chains = 4, 
   parallel_chains = 4,
   iter_warmup = iter/2,
   iter_sampling = iter/2,
   refresh = iter/50
 )

results_path <- here::here(file.path("data", 
                                     iter, 
                                  {str_replace_all(Sys.time(),
                                                      "[- :]",
                                                   "") %>%
                                         str_replace("\\d{2}$",
                                                     "")}))

dir.create(results_path, 
           showWarnings = FALSE, 
           recursive = TRUE)

dcpo_output$save_data_file(dir = results_path,
                           random = FALSE)
dcpo_output$save_output_files(dir = results_path,
                              random = FALSE)
```

```{r dcpo_results, cache=FALSE}
if (!exists("results_path")) {
  latest <- "202305140731"
  results_path <- here::here("data", "1500", latest)
  
  # Define OSF_PAT in .Renviron: https://docs.ropensci.org/osfr/articles/auth
  if (!file.exists(file.path(results_path, paste0("dcpo-", latest, "-1.csv")))) {
    dir.create(results_path, showWarnings = FALSE, recursive = TRUE)
    osf_retrieve_node("XXXXX") %>% 
      osf_ls_files() %>% 
      filter(name == latest) %>% 
      osf_download(path = here::here("data", "1000"))
  }
}

dcpo_output <- as_cmdstan_fit(here::here(results_path,
                                         list.files(results_path, pattern = "csv$")))
 

```

```{r dcpo_results_summary, cache=TRUE}
load(file = here::here("data", "dcpo_input.rda"))
theta_results <- DCPOtools::summarize_dcpo_results(dcpo_input,
                                                   dcpo_output,
                                                   "theta")

res_cy <- nrow(theta_results) %>% 
  scales::comma()

res_c <- theta_results %>% 
  pull(country) %>% 
  unique() %>% 
  length()

save(theta_results, file = here::here("data","theta_results.rda"))
```


```{r cs_plot, fig.cap="TCS Scores, Most Recent Available Year \\label{cs_mry}", fig.height=10, fig.width=8}

n_panes <- 2
axis_text_size <- 10

p1_data <- theta_results %>%
  group_by(country) %>%
  top_n(1, year) %>%
  ungroup() %>%
  arrange(mean) %>%
  transmute(country_year = paste0(country, " (", year, ")") %>% 
              str_replace("’", "'"),
            estimate = mean,
            conf.high = q90,
            conf.low = q10,
            pane = n_panes - (ntile(mean, n_panes) - 1),
            ranked = as.factor(ceiling(row_number())))

p_theta <- ggplot(p1_data,
                  aes(x = estimate, y = ranked)) +
  geom_segment(aes(x = conf.low, xend = conf.high,
                   y = ranked, yend = ranked),
               na.rm = TRUE,
               alpha = .4) +
  geom_point(fill = "black", shape = 21, size = .5, na.rm = TRUE) +
  theme_bw() + theme(legend.position="none",
                     axis.text.x  = element_text(size = axis_text_size,
                                                 angle = 90,
                                                 vjust = .45,
                                                 hjust = .95),
                     axis.text.y  = element_text(size = axis_text_size),
                     axis.title = element_blank(),
                     strip.background = element_blank(), 
                     strip.text = element_blank(),
                     panel.grid.major = element_line(linewidth = .3),
                     panel.grid.minor = element_line(linewidth = .15)) +
  scale_y_discrete(breaks = p1_data$ranked, labels=p1_data$country_year) +
  coord_cartesian(xlim=c(0, 1)) +
  facet_wrap(vars(pane), scales = "free", nrow = 1)


p_theta +
  plot_annotation(caption = "Note: Gray whiskers represent 80% credible intervals.")

bottom5 <- p1_data %>% 
  arrange(ranked) %>% 
  slice(1:5) %>% 
  pull(country_year) %>% 
  str_replace(" \\(.*", "") %>% 
  knitr::combine_words()

```


```{r ts_plots, fig.cap="Tolerance Over Time Within Selected Countries \\label{ts}", fig.height=3.5}
countries <- c("Netherlands", "Sweden", "Ireland", "United States",
               "Argentina", "Mexico", "Taiwan", "Poland", 
               "China", "Ukraine", "Indonesia", "Russia", 
               "Nigeria", "Jamaica", "Armenia", "Uganda")

c_res <- theta_results %>% 
  filter(country %in% countries) %>%
  mutate(country = factor(country, levels = countries))

ggplot(data = c_res, aes(x = year, y = mean)) +
  theme_bw() +
  theme(legend.position = "none") +
  coord_cartesian(xlim = c(1985, 2020), ylim = c(0, 1)) +
  labs(x = NULL, y = "DSD Scores") +
  geom_ribbon(data = c_res, aes(ymin = q10, ymax = q90, linetype=NA), alpha = .25) +
  geom_line(data = c_res) +
  facet_wrap(~as_factor(country), nrow = 2) +
  theme(axis.text.x  = element_text(size=7,
                                    angle = 90,
                                    vjust = .45,
                                    hjust = .95),
        strip.background = element_rect(fill = "white", colour = "white")) +
  plot_annotation(caption = "Note: Countries are ordered by their DSD scores in their most recent\navailable year; gray shading represents 80% credible intervals.")
```


# Appendix A: Survey Items Used to Estimate Macrointerest
National and cross-national surveys have often included questions tapping interest in politics over the past four decades, but the resulting data are both sparse, that is, unavailable for many countries and years, and incomparable, generated by many different survey items.
In all, I identified `r n_items` such survey items that were asked in no fewer than five country-years in countries surveyed at least twice; these items were drawn from `r n_surveys` different survey datasets.
These items are listed in the table below, along with the dispersion ($\alpha$) and difficulty ($\beta$) scores estimated for each from the DCPO model.
Lower values of dispersion indicate questions that better identify publics with a higher level of trust from those with lower.
Items have one less difficulty score than the number of response categories.
Survey dataset codes correspond to those used in the `DCPOtools` R package [@Solt2019]; they appear in decreasing order of country-years contributed.

Together, the survey items in the source data were asked in `r n_countries` different countries in at least two time points over `r n_years` years, `r year_range`, yielding a total of `r n_cyi` country-year-item observations.
The number of items observed in the source data for each country-year is plotted in Figure\nobreakspace{}\ref{obs_by_cy} below.
The estimates of tolerance in country-years with more observed items are more precise.
In country-years with fewer observed items, the estimates rely more heavily on the random-walk prior and are therefore more uncertain, and when there are no observed items, the estimates rely _entirely_ on the random-walk prior and so uncertainty increases still further.

\pagebreak

\noindent Table A1: Indicators Used in the Latent Variable Model of Macrointerest

```{r dcpo_items}
alpha_results <- summarize_dcpo_results(dcpo_input,
                                        dcpo_output = dcpo_output,
                                        pars = "alpha") %>% 
  transmute(item = question,
            dispersion = mean)

beta_results <- summarize_dcpo_results(dcpo_input,
                                       dcpo_output,
                                       "beta") %>% 
  group_by(question) %>% 
  summarize(difficulties0 = paste0(sprintf("%.2f", round(mean, 2)),
                                   collapse = ", ")) %>% 
  mutate(item = question,
         cp = if_else(str_detect(item, "threestate"),
                      2, 
                      as.numeric(str_extract(item, "\\d+")) - 1),
         term = str_glue("(( ?-?[0-9].[0-9][0-9]?,?){{{cp}}})"),
         difficulties = str_extract(difficulties0, 
                                    term) %>%
           str_replace(",$", "") %>% 
           str_trim()) %>% 
  transmute(item, difficulties)
                                    
items_summary <- dcpo_input_raw1 %>%
  dplyr::select(country, year, item, survey) %>%
  distinct() %>%
  separate(survey, c("surv1", "surv2", "surv3"), sep=", ", fill = "left") %>%
  pivot_longer(cols = starts_with("surv"), values_to = "survey") %>%
  filter(!is.na(survey)) %>% 
  group_by(item) %>% 
  mutate(survey = str_extract(survey, "^[a-z]*"),
         all_surveys = paste0(unique(survey), collapse = ", ")) %>% 
  ungroup() %>% 
  distinct(country, year, item, .keep_all = TRUE) %>% 
  group_by(item) %>% 
  mutate(n_cy = n()) %>% 
  ungroup() %>%
  distinct(item, n_cy, all_surveys) %>% 
  left_join(surveys_gm %>%
              select(item, question_text, response_categories) %>%
              distinct(item, .keep_all = TRUE),
            by = "item") %>% 
  left_join(alpha_results, by = "item") %>% 
  left_join(beta_results, by = "item") %>% 
  arrange(-n_cy)
```

```{r dcpo_items_table}  
items_summary %>% 
  transmute(`Survey\nItem\nCode` = item,
            `Country-Years` = as.character(n_cy),
            `Question Text` = str_replace(question_text, "([^(]*)\\(.*", "\\1"),
            `Response Categories` = response_categories,
            `Dispersion` = dispersion,
            `Difficulties`= difficulties,
            `Survey Dataset Codes` = all_surveys) %>% 
  modelsummary::datasummary_df(output = "kableExtra",
                               longtable = TRUE) %>% 
  kableExtra::column_spec(1, width = "7em") %>%
  kableExtra::column_spec(2, width = "4em") %>%
  kableExtra::column_spec(3, width = "13em") %>%
  kableExtra::column_spec(4, width = "16em") %>%
  kableExtra::column_spec(5, width = "4em") %>%
  kableExtra::column_spec(c(6, 7), width = "8em") %>% 
  kableExtra::kable_styling(font_size = 7) %>%
  kableExtra::kable_styling(latex_options = c("repeat_header")) %>%
  kableExtra::kable_styling(latex_options = "striped")
```

\pagebreak
